Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal

### 2020-08-31 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
added manual recipe to build pdf from tex

\begin{abstract} % 250 words

% estimate parameters from the signal itself
% broadband
% Dominant frequency as by-product
% we define several entities in the process, introducing a new measure of discrete curve curvature, hoping to consolidate the terminology in the area

  % background  
  Despite being an elusive concept, mathematically well defined only for artificially generated waves, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information carrying medium in spoken voice and telecommunications, for example.
  intuitively, the temporal envelope can be understood as a smooth function on the same variable as the principal wave, or temporal fine structure, that modulates the signal, responsible for its outer shape. It is implied in this definition that the envelope is non periodic in general and is thus better addressed by specific techniques diverse from those used in the analysis of the periodic, rapidly changing wave it encompasses.
  % motivation
  Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition. Despite of that, a general approach to envelope detection of signals with rich spectral content seems to be lacking, and most methods involve manual intervention, like filter design, based on prior knowledge about the wave under investigation.
  % objective
  In this paper we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
  % methods
  The approach here described draws inspiration from general, computational and differential geometry to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end The concepts of alpha-shapes, concave hulls, and discrete curve curvature are explored. We also define a pulse in the context of a arbitrary digital signal as a means to reduce dimensionality and the complexity of the proposed algorithm.
  % results

  % implications

\end{abstract}

@@

\section{Introduction} % 500 words
% Opener sentence
Despite being ubiquitous in digital signal processing, the literature about envelope detection is very fragmented \parencite{2017LyonsDigital}. Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise contents, in line with the most common usages of those algorithms with artificial signals in the context of analog telecommunications.

@@

\section{Results}
% TODO
Figure \ref{fig:FullFrontiers} illustrates the frontiers obtained by the application of the method to 6 diverse discrete sound waves.
% Revisit purpose and methods
% Overview of results
% Key results in details

@@

>>>>

\begin{abstract} % 250 words
% background  
Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information carrying medium in spoken voice and telecommunications, for example.
intuitively, the temporal envelope can be understood as a smooth function, that modulates the signal, being responsible for its outer shape.
% motivation
Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition. Despite of that, a general approach to envelope detection of signals with rich spectral content seems to be lacking, and most methods involve manual intervention, like filter design or smoothing, based on prior knowledge about the wave under investigation.
% objective
In this paper we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, completely eliminating the necessity of parameter tuning.
% methods
The approach here described draws inspiration from general, computational and differential geometry to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end The concepts of alpha-shapes, concave hulls, and discrete curvature are explored. We also define a pulse in the context of a arbitrary digital signal as a means to reduce dimensionality and the complexity of the proposed algorithm.
% results
We find that the algorithm accurate in the identification of the frontiers of a range of digital sound signal with very diverse characteristics, while providing the dominant frequency of the wave as a by-product. The algorithm also compares favourably with a common detection technique based on the Hilbert Transform.
While the results are satisfactory per se, we define several geometric entities in this work, including a new measure of discrete curvature; we hope to consolidate theory in the area, while providing a solid basis for the development of new techniques, specially in sound synthesis.
\end{abstract}

@@

\section{Introduction} % 500 words
% Opener sentence
Despite being ubiquitous in both analog and digital signal processing, the literature about envelope detection is very fragmented \parencite{2017LyonsDigital}. Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise contents, in line with the most common usages of those algorithms with artificial signals in the context of analog telecommunications; that limitation arises in part by the lack of a mathematical definition for a temporal envelope \parencite{2013Mengempirical}.

@@

\section{Results}
% TODO
% Revisit purpose and methods
% Overview of results
Figure \ref{fig:FullFrontiers} illustrates the frontiers obtained to six diverse discrete sound waves, as well as a detail view of a shorter segment for each wave. All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real world scenarios.

We can see that the frontier is satisfactorily detected in harmonic an inharmonic sounds, and presents robustness in relation to the number of samples and the frequencies of the waves.

% Key results in details

@@

The impact of the normalization of the wave with the frontiers obtained can be seen in figure \ref{fig:Fourier}, which shows the original and normalized wave in the case of a recording of the key 33 of a piano both in the time and frequency domains.

\begin{figure}[ht!]
  \centering
    \includegraphics[width=0.8\linewidth]{08Fourier.pdf}
  \caption{Original and normalized views of a recording of key 33 (F3, 175 Hz) from a grand piano both in time and frequency domains.}
  \label{fig:Fourier}
\end{figure}
